---
title: "Dropcam data analysis - TUV"
subtitle: "Synthesizing dropcam annotations for TUV report"
author: "whitney"
format: html
---

# Prep

This script processes the dscm annotation dataset (provided by DARC) for inclusion in expedition report.  
Data has been prepped using 01_data_prep.qmd and 02_refined_taxa_list.R.  


```{r}
#| label: setup
#| echo: true
#| message: false
#| warning: false
#| results: hide

# for reference later: '#| include: false' hides the code and output from the final report


library(tidyverse)
library(gt)
library(PristineSeasR2)
# library(litedown)  this was in the first script, I dont know if we need it here.

```

```{r}
#| label: configuration
#| echo: false # hides the code from the report

# Define paths
ps_paths <- ps_science_paths()
exp_path <- file.path(ps_paths$expeditions, "TUV-2025")
processed_dir <- file.path(exp_path, "data/primary/processed/dscm")

# Load cleanly processed dataset (from 01_data_prep)
clean_taxa <- read_rds(file.path(processed_dir, "clean_master_data.rds"))

# Load phylogenetic keys for the table sorting
source("scripts/phylo_keys.R") # this has to exist in the scripts folder of the project
```
In prepping the data, we refine our taxonomic richness metrics, by refining our interpretation of what to count as a distinct taxon.  
Without this refinement, we would just be counting how many scientificName entries are fish or inverts.  
But we want to exclude higher-level taxonomic IDs if a lower-level taxon is present, e.g. we don't count both Macrouridae and Coryphaenoides longicirrhus for a species richness of 2.

Here's a step to QA/QC the logic.  
In the process of refining our data, review the table below to see exactly what taxa the code decided to exclude.  
This is a list of every single higher-level taxon that was successfully dropped (like Macrouridae, or Ophiuroidea) because a lower-level ID bumped it out.  
```{r}
#| label: refine-taxa
#| echo: false

# Run the 02_refine_taxa_list.R script to load the specific variables within
# i.e.  refined_taxa_list, true_richness_list, total_taxa_refined, fish_taxa_refined, and invert_taxa_refined

source("scripts/prep/02_refined_taxa_list.R")
```

# Sampling summary

```{r}
#| label: meta-summary-stats
#| echo: false

# Summarize Deployments
deploy_meta <- clean_taxa %>% 
  select(deployment, depth_m, substrate_type) %>% 
  distinct()

total_deps <- nrow(deploy_meta)
mean_depth <- round(mean(deploy_meta$depth_m, na.rm = TRUE), 1)
sd_depth <- round(sd(deploy_meta$depth_m, na.rm = TRUE), 1)
min_depth <- min(deploy_meta$depth_m, na.rm = TRUE)
max_depth <- max(deploy_meta$depth_m, na.rm = TRUE)

n_soft <- sum(str_detect(tolower(deploy_meta$substrate_type), "soft"), na.rm = TRUE)
n_hard <- sum(str_detect(tolower(deploy_meta$substrate_type), "hard"), na.rm = TRUE)

```

We deployed `r total_deps` deep-sea cameras in the Tuvalu EEZ, ranging in depths from `r min_depth` to `r max_depth` m (mean ± SD = `r mean_depth` ± `r sd_depth` m). Substrates represented across deployments were mixed, with about two-thirds of the deployments (n= `r n_hard`) on hard, rocky substrate, and the other third on soft, sandy substrate (n = `r n_soft`).  

** NOTES FOR LATER:  
- refine summary stats by removing deployments that weren't successful or provided any data

```{r}
#| label: taxa-richness
#| echo: false


# Calculate Distinct Taxa
# pull data from 02_refined_taxa_list
 
n_fish <- fish_taxa_refined
n_inverts <- invert_taxa_refined
n_total <- total_taxa_refined
```


Across these deployments, `r n_fish` distinct fish taxa and `r n_inverts` distinct invertebrate taxa were observed, for a total of `r n_total` distinct taxa that were observed on deep sea cameras in Tuvalu.

# Taxa Inventory


SPECIES TABLE
Table of all the taxa observed on deep sea cameras

** NOTES FOR LATER:  
- need to manually refine the taxa in the report, remove redundancies (higher taxa)  

```{r}
#| label: species-table
#| echo: false
#| message: false
#| warning: false

library(gt)

# ===== 1. QUANTITATIVE METRICS ===========

# A. Calculate depth ranges for ALL taxa
taxon_depths <- clean_taxa %>%
  group_by(phylum, class, order, family, scientificName) %>%
  summarise(
    min_d = min(depth_m, na.rm = TRUE),
    max_d = max(depth_m, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Depth_m = if_else(min_d == max_d, 
                      as.character(min_d), 
                      paste0(min_d, " - ", max_d))
  ) %>%
  select(-min_d, -max_d)

# B. Calculate quantitative stats strictly for ATTRACTED taxa
taxon_quant <- clean_taxa %>%
  filter(attracted == "Attracted") %>%
  group_by(phylum, class, order, family, scientificName) %>%
  summarise(
    freq_occ = round((n_distinct(deployment) / total_deps) * 100),
    MaxN = max(individualCount, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Freq_occ = as.character(freq_occ),
    MaxN = as.character(MaxN)
  ) %>%
  select(-freq_occ)

# C. Extract Tentative IDs (Collapsed if there are multiple; formatting with italics before collapsing)
tentative_list <- clean_taxa %>%
  filter(!is.na(tentativeId)) %>%
  mutate(
    # Use the exact same logic to identify tentative species vs genera
    t_is_species = str_detect(tentativeId, " "),
    t_is_genus = !t_is_species & family != "" & tentativeId != family,
    
    # Apply markdown italics 
    fmt_tentativeId = case_when(
      t_is_species ~ paste0("*", tentativeId, "*"),
      # If you want tentative genera to ALSO have " sp.", use paste0("*", tentativeId, "* sp.") below instead:
      t_is_genus ~ paste0("*", tentativeId, "* sp."), 
      TRUE ~ tentativeId
    )
  ) %>%
  group_by(phylum, class, order, family, scientificName) %>%
  summarise(
    # Collapse the newly formatted IDs!
    t_ids = paste(unique(fmt_tentativeId), collapse = ", "),
    .groups = "drop"
  )

# ====== 2. DATA MERGING & FORMATTING ========================================

table_data <- taxon_depths %>%
  left_join(taxon_quant, by = c("phylum", "class", "order", "family", "scientificName")) %>%
  left_join(tentative_list, by = c("phylum", "class", "order", "family", "scientificName")) %>%
  
  mutate(
    # Apply phylogenetic sorting rules
    phylum = factor(phylum, levels = phylo_phylum),
    class = fct_relevel(factor(class), phylo_class),
    order = fct_relevel(factor(order), phylo_order),
    family = fct_relevel(factor(family), phylo_family)
  ) %>%
  arrange(phylum, class, order, family, scientificName) %>%
  
  # Fill NA values with blanks
  mutate(across(everything(), as.character)) %>% 
  mutate(across(everything(), ~replace_na(.x, ""))) %>%
  
  # Text Formatting Logic (Italics & Tentative IDs via Markdown)
  mutate(
    # Identify taxonomy levels based on naming conventions
    is_species = str_detect(scientificName, " "),
    is_genus = !is_species & family != "" & scientificName != family,
    
    # Apply base italics and " sp."
    base_name = case_when(
      is_species ~ paste0("*", scientificName, "*"),
      is_genus ~ paste0("*", scientificName, "* sp."),
      TRUE ~ scientificName
    ),
    
    # If a tentative ID exists, append it with a line break <br>
    scientificName = if_else(
      t_ids != "",
      paste0(base_name, "<br>(Tentative: ", t_ids, ")"),
      base_name
    )
  ) %>%
  # Clean up our temporary logic columns
  select(-is_species, -is_genus, -base_name, -t_ids)

# ====== 3. GENERATE TABLE ==================================================

table_data %>%
  gt() %>%
  tab_header(
    title = "Taxonomic Inventory of Deep-Sea Camera Observations",
    subtitle = "Tuvalu"
  ) %>%
  cols_label(
    phylum = "Phylum",
    class = "Class",
    order = "Order",
    family = "Family",
    scientificName = "Lowest Taxonomic ID",
    Depth_m = "Depth (m)",
    Freq_occ = "Freq. of Occ. (%)",
    MaxN = "MaxN"
  ) %>%
  # Center align math columns
  cols_align(
    align = "center",
    columns = c(Depth_m, Freq_occ, MaxN)
  ) %>%
  # Set tight widths
  cols_width(
    Depth_m ~ px(140),  
    Freq_occ ~ px(90),  
    MaxN ~ px(80),
    # Bumped scientificName slightly to accommodate long tentative text
    scientificName ~ px(220) 
  ) %>%
  # TELL GT TO READ THE ASTERISKS AND <br> AS MARKDOWN/HTML
  fmt_markdown(columns = scientificName) %>%
  # Prevent depth column from wrapping
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = Depth_m)
  ) %>%
  opt_row_striping() %>%
  tab_options(
    data_row.padding = px(5),
    table.font.size = px(12)
  )
```