---
title: "Dropcam data analysis - TUV"
subtitle: "Synthesizing dropcam annotations for TUV report"
author: "whitney"
format: html
---

# Prep

This script processes the dscm annotation dataset (provided by DARC) for inclusion in expedition report.  
Data has been prepped using 01_data_prep.qmd and 02_refined_taxa_list.R.  


```{r}
#| label: setup
#| echo: true
#| message: false
#| warning: false
#| results: hide

# for reference later: '#| include: false' hides the code and output from the final report


library(tidyverse)
library(gt)
library(PristineSeasR2)
# library(litedown)  this was in the first script, I dont know if we need it here.

```

```{r}
#| label: configuration
#| echo: false # hides the code from the report

# Define paths
ps_paths <- ps_science_paths()
exp_path <- file.path(ps_paths$expeditions, "TUV-2025")
processed_dir <- file.path(exp_path, "data/primary/processed/dscm")

# Load cleanly processed dataset (from 01_data_prep)
clean_taxa <- read_rds(file.path(processed_dir, "clean_master_data.rds"))

# Load phylogenetic keys for the table sorting
source("scripts/phylo_keys.R") # this has to exist in the scripts folder of the project
```
In prepping the data, we refine our taxonomic richness metrics, by refining our interpretation of what to count as a distinct taxon.  
Without this refinement, we would just be counting how many scientificName entries are fish or inverts.  
But we want to exclude higher-level taxonomic IDs if a lower-level taxon is present, e.g. we don't count both Macrouridae and Coryphaenoides longicirrhus for a species richness of 2.

Here's a step to QA/QC the logic.  
In the process of refining our data, review the table below to see exactly what taxa the code decided to exclude.  
This is a list of every single higher-level taxon that was successfully dropped (like Macrouridae, or Ophiuroidea) because a lower-level ID bumped it out.  
```{r}
#| label: refine-taxa
#| echo: false

# Run the 02_refine_taxa_list.R script to load the specific variables within
# i.e.  refined_taxa_list, true_richness_list, total_taxa_refined, fish_taxa_refined, and invert_taxa_refined

source("scripts/prep/02_refined_taxa_list.R")
```

# Sampling summary

```{r}
#| label: meta-summary-stats
#| echo: false

# Summarize Deployments
deploy_meta <- clean_taxa %>% 
  select(deployment, depth_m, substrate_type, valid_for_freq) %>% 
  distinct()

total_deps <- nrow(deploy_meta)
valid_freq_deps <- deploy_meta %>% filter(valid_for_freq == TRUE) %>% nrow() # Accounts for excluded deployments (that were incomplete); Calculate the denominator specifically for frequency of occ. math
mean_depth <- round(mean(deploy_meta$depth_m, na.rm = TRUE), 1)
sd_depth <- round(sd(deploy_meta$depth_m, na.rm = TRUE), 1)
min_depth <- min(deploy_meta$depth_m, na.rm = TRUE)
max_depth <- max(deploy_meta$depth_m, na.rm = TRUE)

n_soft <- sum(str_detect(tolower(deploy_meta$substrate_type), "soft"), na.rm = TRUE)
n_hard <- sum(str_detect(tolower(deploy_meta$substrate_type), "hard"), na.rm = TRUE)

```

We deployed `r total_deps` deep-sea cameras in the Tuvalu EEZ, ranging in depths from `r min_depth` to `r max_depth` m (mean ± SD = `r mean_depth` ± `r sd_depth` m). Substrates represented across deployments were mixed, with about two-thirds of the deployments (n= `r n_hard`) on hard, rocky substrate, and the other third on soft, sandy substrate (n = `r n_soft`).  

```{r}
#| label: fig-depth-profile
#| fig-cap: "Distribution of camera deployments across depth. Stacked colors indicate deployments included in quantitative frequency analyses versus partial deployments used only for taxon presence and MaxN."
#| echo: false
#| message: false
#| warning: false

library(ggplot2)
library(dplyr)

# 1. Prep the data: Make the TRUE/FALSE flag human-readable for the legend

plot_data <- deploy_meta %>%
  # Drop any rows where depth might be missing just in case
  filter(!is.na(depth_m)) %>% 
  mutate(
    effort_status = case_when(
      valid_for_freq == TRUE ~ "Full Deployment (All Metrics)",
      valid_for_freq == FALSE ~ "Partial Deployment (Presence/MaxN Only)",
      is.na(valid_for_freq) ~ "Partial Deployment (Presence/MaxN Only)", # Catches any stray NAs
      TRUE ~ "Partial Deployment (Presence/MaxN Only)"
    ),
    # Factor to force the order in the legend and the stack
    effort_status = factor(effort_status, levels = c("Full Deployment (All Metrics)", "Partial Deployment (Presence/MaxN Only)"))
  )

# 2. Build the ggplot
# Dynamically calculate the bottom of the y-axis, rounding up to the nearest 100
max_depth <- ceiling(max(plot_data$depth_m, na.rm = TRUE) / 100) * 100

ggplot(plot_data, aes(y = depth_m, fill = effort_status)) +
  # Using a histogram mapped directly to the y-axis makes it horizontal
  geom_histogram(
    binwidth = 100, # Change binwidth here
    boundary = 0, 
    color = "black", 
    linewidth = 0.3,
    position = position_stack(reverse = TRUE) # <--- fixes order of stacks, with full deployments first
  ) +
  
  # Y-axis inverted so surface is at the top, forced to start exactly at 0 and end at max_depth
  # Labels at 500m, minor gridlines at 200m
 scale_y_reverse(
    limits = c(0, max_depth),
    breaks = seq(0, max_depth, by = 500),       # Text labels every 500m
    minor_breaks = seq(0, max_depth, by = 100), # marks every 100m
    expand = expansion(add = c(100, 0)), # starts at 0, pads 200m at bottom
  # This magic argument forces the minor_breaks to draw physical tick marks!
    guide = guide_axis(minor.ticks = TRUE) 
  ) +
  
  # X-axis with whole numbers for deployment counts
  scale_x_continuous(
    limits = c(0, 8), # x-axis to 8
    breaks = seq(0, 8, by = 1),
    expand = expansion(mult = c(0, 0)) # Keeps 0 flush, no padding on right
  ) +
  
  scale_fill_manual(values = c(
    "Full Deployment (All Metrics)" = "#2c7fb8",
    "Partial Deployment (Presence/MaxN Only)"  = "#9ecae1"   # A complementary light blue 
  )) +
  
  # Clean up labels
  labs(
    y = "Depth (m)",
    x = "Number of Deployments",
    fill = NULL # no title
  ) +
  
  # Use a clean theme perfect for reports
  theme_minimal(base_size = 12) +
  theme(
    # --- LEGEND PLACEMENT ---
    # c(1, 1) is the absolute top-right corner of the plotting area
    legend.position = c(0.98, 0.98), 
    legend.justification = c(1, 1), # Anchors the legend box to the top-right coordinate
    legend.direction = "vertical",  # Stacks the keys vertically
    legend.background = element_rect(fill = "white", color = NA), # Adds a tiny invisible white shield behind it
    legend.text = element_text(size = 8), # font size
    
    # --- AXIS TITLE SPACING ---
    # margin(t, r, b, l) - Adds 15 points of space to the Right side of the Y-axis title
    axis.title.y = element_text(margin = margin(r = 15), color = "black"),
    axis.title.x = element_text(color = "black"),   # gridline controls
    
    panel.grid = element_blank(), # Turn off ALL background gridlines
    axis.text = element_text(color = "black"),
    axis.line = element_line(color = "black"),
    
    # physical tick marks on (theme_minimal usually hides them)
    axis.ticks = element_line(color = "black"),
    
    # control tick lengths (Minor ticks will automatically be half the size)
    axis.ticks.length = unit(0.2, "cm"),
    axis.minor.ticks.length = rel(0.3) 
  )
```


```{r}
#| label: taxa-richness
#| echo: false


# Calculate Distinct Taxa
# pull data from 02_refined_taxa_list
 
n_fish <- fish_taxa_refined
n_inverts <- invert_taxa_refined
n_total <- total_taxa_refined
```


Across these deployments, `r n_fish` distinct fish taxa and `r n_inverts` distinct invertebrate taxa were observed, for a total of `r n_total` distinct taxa that were observed on deep sea cameras in Tuvalu.

# Taxa Inventory  

Worth noting this doesn't account for distinct morphospecies. This dataset didn't have any that would change the biodiversity metrics, but for future datasets, will need to check for differentiations by morphospecies.  

SPECIES TABLE  
Table of all the taxa observed on deep sea cameras
```{r}
#| label: species-table
#| echo: false
#| message: false
#| warning: false

library(gt)

# List of taxa to manually exclude from the visual table
# Build this as you go; just keep adding names here separated by commas.
table_exclusions <- c(
  "Animalia",
  "Myxinidae",
  "Chordata",
  "Actinopterygii",
  "Actinopteri",
  "Lutjanidae",
  "Epigonidae",
  "Arthropoda"
)


# ===== 1. QUANTITATIVE METRICS ===========

# A. Calculate depth ranges for ALL taxa
taxon_depths <- clean_taxa %>%
  group_by(phylum, class, order, family, scientificName) %>%
  summarise(
    min_d = min(depth_m, na.rm = TRUE),
    max_d = max(depth_m, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Depth_m = if_else(min_d == max_d, 
                      as.character(min_d), 
                      paste0(min_d, " - ", max_d))
  ) %>%
  select(-min_d, -max_d)

# B. Calculate quantitative stats strictly for ATTRACTED taxa
taxon_quant <- clean_taxa %>%
  filter(attracted == "Attracted") %>%
  group_by(phylum, class, order, family, scientificName) %>%
  summarise(
    # FREQ: Only count deployments where valid_for_freq is TRUE, divide by valid_freq_deps
    freq_occ = round((n_distinct(deployment[valid_for_freq == TRUE]) / valid_freq_deps) * 100),
    # MAXN: Keep using all deployments, no changes
    MaxN = max(individualCount, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Freq_occ = as.character(freq_occ),
    MaxN = as.character(MaxN)
  ) %>%
  select(-freq_occ)

# C. Extract Tentative IDs (Collapsed if there are multiple; formatting with italics before collapsing)
tentative_list <- clean_taxa %>%
  filter(!is.na(tentativeId)) %>%
  mutate(
    # Use the exact same logic to identify tentative species vs genera
    t_is_species = str_detect(tentativeId, " "),
    t_is_genus = !t_is_species & family != "" & tentativeId != family,
    
    # Apply markdown italics 
    fmt_tentativeId = case_when(
      t_is_species ~ paste0("*", tentativeId, "*"),
      # If you want tentative genera to ALSO have " sp.", use paste0("*", tentativeId, "* sp.") below instead:
      t_is_genus ~ paste0("*", tentativeId, "* sp."), 
      TRUE ~ tentativeId
    )
  ) %>%
  group_by(phylum, class, order, family, scientificName) %>%
  summarise(
    # Collapse the newly formatted IDs!
    t_ids = paste(unique(fmt_tentativeId), collapse = ", "),
    .groups = "drop"
  )

# ====== 2. DATA MERGING & FORMATTING ========================================

table_data <- taxon_depths %>%
  left_join(taxon_quant, by = c("phylum", "class", "order", "family", "scientificName")) %>%
  left_join(tentative_list, by = c("phylum", "class", "order", "family", "scientificName")) %>%
  
  # manual exclusions
  filter(!scientificName %in% table_exclusions) %>%
  
  mutate(
    # Apply phylogenetic sorting rules
    phylum = factor(phylum, levels = phylo_phylum),
    class = fct_relevel(factor(class), phylo_class),
    order = fct_relevel(factor(order), phylo_order),
    family = fct_relevel(factor(family), phylo_family)
  ) %>%
  arrange(phylum, class, order, family, scientificName) %>%
  
  # Fill NA values with blanks
  mutate(across(everything(), as.character)) %>% 
  mutate(across(everything(), ~replace_na(.x, ""))) %>%
  
  # Text Formatting Logic (Italics & Tentative IDs via Markdown)
  mutate(
    # Identify taxonomy levels based on naming conventions
    is_species = str_detect(scientificName, " "),
    is_genus = !is_species & family != "" & scientificName != family,
    
    # Apply base italics and " sp."
    base_name = case_when(
      is_species ~ paste0("*", scientificName, "*"),
      is_genus ~ paste0("*", scientificName, "* sp."),
      TRUE ~ scientificName
    ),
    
    # If a tentative ID exists, append it with a line break <br>
    scientificName = if_else(
      t_ids != "",
      paste0(base_name, "<br>(Tentative: ", t_ids, ")"),
      base_name
    )
  ) %>%
  # Clean up our temporary logic columns
  select(-is_species, -is_genus, -base_name, -t_ids)

# ====== 3. GENERATE TABLE ==================================================

table_data %>%
  gt() %>%
  tab_header(
    title = "Taxonomic Inventory of Deep-Sea Camera Observations",
    subtitle = "Tuvalu"
  ) %>%
  cols_label(
    phylum = "Phylum",
    class = "Class",
    order = "Order",
    family = "Family",
    scientificName = "Lowest Taxonomic ID",
    Depth_m = "Depth (m)",
    Freq_occ = "Freq. of Occ. (%)",
    MaxN = "MaxN"
  ) %>%
  # Center align math columns
  cols_align(
    align = "center",
    columns = c(Depth_m, Freq_occ, MaxN)
  ) %>%
  # Set tight widths
  cols_width(
    Depth_m ~ px(140),  
    Freq_occ ~ px(90),  
    MaxN ~ px(80),
    # Bumped scientificName slightly to accommodate long tentative text
    scientificName ~ px(220) 
  ) %>%
  # TELL GT TO READ THE ASTERISKS AND <br> AS MARKDOWN/HTML
  fmt_markdown(columns = scientificName) %>%
  # Prevent depth column from wrapping
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = Depth_m)
  ) %>%
  opt_row_striping() %>%
  tab_options(
    data_row.padding = px(5),
    table.font.size = px(12)
  )
```


# Exploratory taxa summaries

## Shark richness, Fish families, Abundant taxa  

```{r}
#| label: explore-taxa-interest
#| echo: false
#| message: false
#| warning: false

# 1. Distinct Sharks and Chimaeras
# (Filtering true_richness_list by the Classes we defined in our phylo_keys)
sharks_chimaeras <- true_richness_list %>%
  filter(class %in% c("Elasmobranchii", "Holocephali")) %>%
  nrow()

# Calculate distinct families of bony fish (excluding sharks, chimaeras, hagfish, and pelagic tunicates)
bony_fish_families_count <- clean_taxa %>%
  filter(
    phylum == "Chordata", 
    !class %in% c("Elasmobranchii", "Holocephali", "Myxini", "Appendicularia"),
    family != ""
  ) %>%
  summarise(n = n_distinct(family)) %>%
  pull(n)

cat("--- DISTINCT FISH GROUPS ---\n")
cat("Total non-redundant shark & chimaera taxa:", sharks_chimaeras, "\n")
cat("Total distinct bony fish families:", bony_fish_families_count, "\n\n")

# Calculate phyla
phylum_count <- clean_taxa %>%
  summarise(n = n_distinct(phylum)) %>%
  pull(n)

cat("--- DISTINCT PHYLA ---\n")
cat("Total number of phyla:", phylum_count, "\n")

# 2. Distinct Chordate Families & Frequencies of Occurrence
chordate_families_count <- clean_taxa %>%
  filter(phylum == "Chordata", family != "") %>%
  summarise(n = n_distinct(family)) %>%
  pull(n)

all_chordate_families <- clean_taxa %>%
  # Add valid_for_freq == TRUE to the filter, so incomplete deployments aren't included in freq. of occ.
  filter(phylum == "Chordata", family != "", valid_for_freq == TRUE) %>%
  group_by(family) %>%
  # Divide by valid_freq_deps
  summarise(freq_occ = round((n_distinct(deployment) / valid_freq_deps) * 100, 1)) %>%
  arrange(desc(freq_occ))

cat("--- CHORDATE FAMILIES ---\n")
cat("Total distinct chordate families:", chordate_families_count, "\n")
cat("Frequency of occurrence for ALL chordate families:\n")
print(all_chordate_families, n = Inf) 
cat("\n")

# 2b. Distinct invert orders & Frequencies of Occurrence
invert_order_count <- clean_taxa %>%
  filter(phylum != "Chordata", order != "") %>%
  summarise(n = n_distinct(order)) %>%
  pull(n)

all_invert_orders <- clean_taxa %>%
  # Add valid_for_freq == TRUE to the filter, so incomplete deployments aren't included in freq. of occ.
  filter(phylum != "Chordata", order != "", valid_for_freq == TRUE) %>%
  group_by(order) %>%
  # Divide by valid_freq_deps
  summarise(freq_occ = round((n_distinct(deployment) / valid_freq_deps) * 100, 1)) %>%
  arrange(desc(freq_occ))

cat("--- INVERT ORDERS ---\n")
cat("Total distinct invert orders:", invert_order_count, "\n")
cat("Frequency of occurrence for ALL invert orders:\n")
print(all_invert_orders, n = Inf) 
cat("\n")

# 3. Highest MaxN (Abundance) - for Fish and Inverts
top_maxn_fish <- clean_taxa %>%
  filter(!is.na(individualCount), phylum == "Chordata") %>%
  group_by(scientificName) %>%
  summarise(MaxN = max(individualCount, na.rm = TRUE), .groups = "drop") %>%
  # Grab top 10
  slice_max(order_by = MaxN, n = 10, with_ties = FALSE) %>%
  arrange(desc(MaxN))

top_maxn_inverts <- clean_taxa %>%
  filter(!is.na(individualCount), phylum != "Chordata") %>%
  group_by(scientificName) %>%
  summarise(MaxN = max(individualCount, na.rm = TRUE), .groups = "drop") %>%
  # Grab top 10
  slice_max(order_by = MaxN, n = 10, with_ties = FALSE) %>%
  arrange(desc(MaxN))

cat("--- TOP 10 MOST ABUNDANT FISHES (Highest MaxN) ---\n")
print(top_maxn_fish)
cat("\n")

cat("--- TOP 10 MOST ABUNDANT INVERTEBRATES (Highest MaxN) ---\n")
print(top_maxn_inverts)
```
# Characteristic fauna by depth zone  
For this chunk, we are going to define your depth bins using "case_when()". 

**Important ecological math note:** When calculating the frequency of occurrence for a depth zone, we don't divide by the *total* number of deployments on the expedition. We divide by the number of deployments *inside that specific depth bin*. (e.g., If a fish is seen on 5 deep deployments, and there were only 6 deep deployments total, it has an 83% frequency for that zone, making it highly characteristic!)
```{r}
#| label: explore-depth-zones
#| echo: false
#| message: false
#| warning: false

# 1. Define Depth Zones and calculate deployments per zone
zone_deps <- deploy_meta %>%
  mutate(
    depth_zone = case_when(
      depth_m < 700 ~ "<700 m",
      depth_m >= 700 & depth_m <= 1500 ~ "700-1500 m",
      depth_m > 1500 ~ ">1500 m"
    ),
    depth_zone = factor(depth_zone, levels = c("<700 m", "700-1500 m", ">1500 m"))
  ) %>%
  filter(!is.na(depth_zone)) %>%
  group_by(depth_zone) %>%
  summarise(
    total_zone_deps = n(),
    # NEW: Count how many VALID deployments are in each zone
    valid_zone_deps = sum(valid_for_freq == TRUE), 
    .groups = "drop"
  )

# Merge zone definitions and fauna groups into the main data
zone_data <- clean_taxa %>%
  mutate(
    depth_zone = case_when(
      depth_m < 700 ~ "<700 m",
      depth_m >= 700 & depth_m <= 1500 ~ "700-1500 m",
      depth_m > 1500 ~ ">1500 m"
    ),
    depth_zone = factor(depth_zone, levels = c("<700 m", "700-1500 m", ">1500 m")),
    fauna_group = if_else(phylum == "Chordata", "Fish", "Invertebrate")
  ) %>%
  left_join(zone_deps, by = "depth_zone") %>%
  filter(!is.na(depth_zone)) # Drop rows without depth data

# =========================================================
# ROUND A: Characteristic Taxa by Lowest ID (scientificName)
# =========================================================
zone_sp_freq <- zone_data %>%
  filter(scientificName != "", valid_for_freq == TRUE) %>%
  group_by(depth_zone, fauna_group, scientificName) %>%
  summarise(
    occ = n_distinct(deployment),
    # Divide occurrences by the number of deployments IN THAT ZONE
    freq_occ = round((n_distinct(deployment) / first(valid_zone_deps)) * 100, 1),
    .groups = "drop"
  ) %>%
  group_by(depth_zone, fauna_group) %>%
  # Grab top 10 for EACH zone and fauna group
  slice_max(order_by = freq_occ, n = 10, with_ties = FALSE) %>%
  arrange(depth_zone, desc(freq_occ)) %>%
  ungroup()

cat("=== ROUND A: TOP 10 TAXA BY DEPTH (Lowest ID) ===\n\n")

cat("--- FISHES ---\n")
print(zone_sp_freq %>% filter(fauna_group == "Fish") %>% select(-fauna_group), n = Inf)
cat("\n")

cat("--- INVERTEBRATES ---\n")
print(zone_sp_freq %>% filter(fauna_group == "Invertebrate") %>% select(-fauna_group), n = Inf)
cat("\n")

# =========================================================
# ROUND B: Characteristic Taxa by Fish Family / Invert Order
# =========================================================
zone_higher_freq <- zone_data %>%
  mutate(
    # Create a dynamic grouping column: Family for Fish, Order for Inverts
    taxa_group = if_else(fauna_group == "Fish", family, order)
  ) %>%
  # Filter out blanks just in case a record lacks family/order
  filter(taxa_group != "", valid_for_freq == TRUE) %>%
  group_by(depth_zone, fauna_group, taxa_group) %>%
  summarise(
    occ = n_distinct(deployment),
    freq_occ = round((n_distinct(deployment) / first(valid_zone_deps)) * 100, 1),
    .groups = "drop"
  ) %>%
  group_by(depth_zone, fauna_group) %>%
  # Grab top 10 for EACH zone and fauna group
  slice_max(order_by = freq_occ, n = 10, with_ties = FALSE) %>%
  arrange(depth_zone, desc(freq_occ)) %>%
  ungroup()

cat("=== ROUND B: TOP 10 HIGHER TAXA BY DEPTH (Fish Family / Invert Order) ===\n\n")

cat("--- FISHES (By Family) ---\n")
print(zone_higher_freq %>% filter(fauna_group == "Fish") %>% select(-fauna_group), n = Inf)
cat("\n")

cat("--- INVERTEBRATES (By Order) ---\n")
print(zone_higher_freq %>% filter(fauna_group == "Invertebrate") %>% select(-fauna_group), n = Inf)
```

If you run those chunks interactively (the green play button), RStudio will print out neat, text-based summaries directly below the chunk. This makes it incredibly easy to scan the outputs and start writing paragraphs like: 

*"The deep-sea community was dominated by... with the family Macrouridae occurring in X% of deployments. In the mesopelagic zone (<700m), the characteristic fauna shifted to..."*

The deep zone is pretty distinct. Let's look at whether there are taxa that occurred here that also occurred elsewhere. 
```{r}
#| label: explore-deep-taxa
#| echo: false
#| message: false
#| warning: false

# 1. Identify all unique taxa seen deeper than 1900 m
deep_taxa_list <- clean_taxa %>%
  filter(depth_m > 1900, scientificName != "") %>%
  distinct(scientificName) %>%
  pull(scientificName)

# 2. Track those specific taxa across the ENTIRE dataset
deep_taxa_overlap <- clean_taxa %>%
  filter(scientificName %in% deep_taxa_list) %>%
  group_by(phylum, family, scientificName) %>%
  summarise(
    min_depth = min(depth_m, na.rm = TRUE),
    max_depth = max(depth_m, na.rm = TRUE),
    n_deep_deps = n_distinct(deployment[depth_m > 1900]),
    n_shallow_deps = n_distinct(deployment[depth_m <= 1900]),
    .groups = "drop"
  ) %>%
  mutate(
    # Categorize them based on whether they ever showed up shallow
    distribution = if_else(n_shallow_deps == 0, "Exclusive to >1900m", "Shared with Shallower Depths")
  ) %>%
  arrange(distribution, min_depth, desc(n_deep_deps))

# 3. Print the results cleanly
cat("=== FAUNAL DISTINCTNESS: TAXA OBSERVED > 1900m ===\n\n")

cat(paste("Total unique taxa observed > 1900m:", length(deep_taxa_list), "\n\n"))

cat("--- EXCLUSIVE TO >1900m ---\n")
exclusive_taxa <- deep_taxa_overlap %>% filter(distribution == "Exclusive to >1900m")
if(nrow(exclusive_taxa) > 0) {
  print(exclusive_taxa %>% select(family, scientificName, n_deep_deps, min_depth, max_depth), n = Inf)
} else {
  cat("None. All deep taxa were also found shallower.\n")
}
cat("\n")

cat("--- SHARED WITH SHALLOWER DEPTHS ---\n")
shared_taxa <- deep_taxa_overlap %>% filter(distribution == "Shared with Shallower Depths")
if(nrow(shared_taxa) > 0) {
  print(shared_taxa %>% select(family, scientificName, n_deep_deps, n_shallow_deps, min_depth, max_depth), n = Inf)
} else {
  cat("None. All deep taxa are entirely restricted to >1900m.\n")
}
```
