---
title: "TUV dscm data prep"
subtitle: "Synthesizing dropcam annotations for TUV report"
author: "whitney"
format: html
---

# Overview

This script imports raw Tator annotations and metadata, and merges and cleans data from the TUV 2025 expedition.


```{r}
#| label: setup
#| echo: true
#| message: false
#| warning: false
#| results: hide

library(tidyverse)
library(readxl)
library(gt)
library(PristineSeasR2)
library(litedown)

ps_paths <- ps_science_paths()
exp_path <- file.path(ps_paths$expeditions, "TUV-2025")
```

** NOTES FOR LATER:  
- refine manual exclusions (check contradictory attracted/not attracted)
```{r}
#| label: load data
#| include: false # hides the code and output from the final report


# Define data path
excel_file <- file.path(exp_path, "data/primary/raw/dscm/TUV_DOEX0112_dscm_Data_Summary_v2.xlsx")

# 1. Load datasets
tator_data <- read_excel(excel_file, sheet = "Tator_Data_Summary")

# Skip  first row (CATAMI header group name)
metadata <- read_excel(excel_file, sheet = "Field_Log_Metadata", skip = 1)

# 2. Select columns of interest
metadata_subset <- metadata %>%
  select(Deployment, `Depth (m)`, Location, `Substrate (Hard/Soft)`) %>%
  rename(deployment = Deployment)

# 3. Merge metadata with annotation data
master_data <- tator_data %>%
  select(-`depth(m)`) %>% # Remove original depth column to avoid duplicates
  left_join(metadata_subset, by = "deployment") %>%
  rename(depth_m = `Depth (m)`,
         substrate_type = `Substrate (Hard/Soft)`,
         location = Location)

# ==============================================================================
# APPLY MANUAL EXCLUSIONS (QA/QC)
# ==============================================================================
exclusion_file <- file.path(exp_path, "data/primary/raw/dscm/manual_exclusions.csv")

if (file.exists(exclusion_file)) {
  # Read in the exclusions file
  exclusions <- read_csv(exclusion_file, show_col_types = FALSE)
  
  # Store row count before exclusion for a helpful message
  rows_before <- nrow(master_data)
  
  # Drop the matching records from master_data
  master_data <- master_data %>%
    anti_join(exclusions, by = c("deployment", "scientificName"))
  
  rows_after <- nrow(master_data)
  message(paste("Applied manual exclusions. Dropped", (rows_before - rows_after), "records."))
  
} else {
  message("No manual_exclusions.csv file found. Proceeding with all records.")
}

# 4. Split data into general observations and quantitative attracted scavengers
species_list_data <- master_data 

quant_data <- master_data %>%
  filter(attracted == "Attracted")


message("Data prep complete!")
```
Basic summary: calculates all the values to reference in report text.  
** NOTES FOR LATER:  
- need to refine code to remove unsuccessful (i.e. no data) deployments from summary stats  
- need to refine distinct taxa counts, i.e. remove higher-level taxonomies if lower-level are also present in the dataset.  
```{r}
#| label: meta-summary-stats
#| include: false

# 1. Deployment Metadata Summaries (1 row per deployment)
deploy_meta <- species_list_data %>%
  select(deployment, depth_m, substrate_type) %>%
  distinct()

total_deps <- nrow(deploy_meta)
mean_depth <- round(mean(deploy_meta$depth_m, na.rm = TRUE), 1)
sd_depth <- round(sd(deploy_meta$depth_m, na.rm = TRUE), 1)
min_depth <- min(deploy_meta$depth_m, na.rm = TRUE)
max_depth <- max(deploy_meta$depth_m, na.rm = TRUE)

n_soft <- sum(str_detect(tolower(deploy_meta$substrate_type), "soft"), na.rm = TRUE)
n_hard <- sum(str_detect(tolower(deploy_meta$substrate_type), "hard"), na.rm = TRUE)

# Broad taxa summaries
clean_taxa <- species_list_data %>%
  filter(!is.na(scientificName)) %>%
  mutate(fauna_group = if_else(phylum == "Chordata", "Fish", "Invertebrate"))

# Get distinct counts
taxa_counts <- clean_taxa %>%
  group_by(fauna_group) %>%
  summarise(total = n_distinct(scientificName))

n_fish <- taxa_counts %>% filter(fauna_group == "Fish") %>% pull(total)
n_inverts <- taxa_counts %>% filter(fauna_group == "Invertebrate") %>% pull(total)


```

We deployed `r total_deps` deep-sea cameras in the Tuvalu EEZ, ranging in depths from `r min_depth` to `r max_depth` m (mean ± SD = `r mean_depth` ± `r sd_depth` m). Substrates represented across deployments were mixed, with about two-thirds of the deployments (n= `r n_hard`) on hard, rocky substrate, and the other third on soft, sandy substrate (n = `r n_soft`). Across these deployments, `r n_fish` distinct fish taxa and `r n_inverts` distinct invertebrate taxa were observed.

SPECIES TABLE
Table of all the taxa observed on deep sea cameras

** NOTES FOR LATER:  
- need to refine the taxa, remove redundancies (higher taxa)  
- create phylo_keys.R script to 'source()' instead of defining vectors. 

```{r}
#| label: species-table
#| echo: false
#| message: false
#| warning: false

library(gt)

# ==============================================================================
# 1. CONFIGURATION: PHYLOGENETIC HIERARCHIES
# ==============================================================================
phylo_phyla <- c(
  "Foraminifera", "Microsporidia", "Porifera", "Cnidaria", 
  "Ctenophora", "Nemertea", "Annelida", "Mollusca", 
  "Arthropoda", "Echinodermata", "Chaetognatha", "Chordata"
)

phylo_classes <- c(
  "Hexactinellida", "Hexacorallia", "Hydrozoa", "Octocorallia", "Scyphozoa",
  "Tentaculata", "Polychaeta", "Copepoda", "Malacostraca", "Cephalopoda",
  "Gastropoda", "Asteroidea", "Crinoidea", "Echinoidea", "Holothuroidea",
  "Ophiuroidea", "Appendicularia", "Myxini", "Holocephali", "Elasmobranchii", "Teleostei"
)

phylo_orders <- c(
  "Actiniaria", "Anthoathecata", "Narcomedusae", "Trachymedusae", 
  "Siphonophora", "Siphonophorae", "Malacalcyonacea", "Scleralcyonacea", 
  "Coronatae", "Cydippida", "Terebellida", "Amphipoda", "Decapoda", 
  "Euphausiacea", "Mysida", "Nautilida", "Buccinida", "Paxillosida", 
  "Valvatida", "Comatulida", "Isocrinida", "Aspidodiadematoida", 
  "Cidaroida", "Diadematoida", "Echinothurioida", "Elasipodida", 
  "Euryalida", "Chimaeriformes", "Aulopiformes", "Lampriformes", 
  "Ophidiiformes", "Perciformes", "Siluriformes", "Stomiiformes", "Trachichthyiformes"
)

phylo_families <- c(
  "Stylasteridae", "Aeginidae", "Rhopalonematidae", "Acanthogorgiidae", 
  "Balticinidae", "Primnoidae", "Periphyllidae", "Acrocirridae", 
  "Eurytheneidae", "Acanthephyridae", "Aristeidae", "Calappidae", 
  "Crangonidae", "Eumunididae", "Geryonidae", "Homolidae", "Lithodidae", 
  "Lyreididae", "Munididae", "Munidopsidae", "Nematocarcinidae", 
  "Nephropidae", "Oregoniidae", "Pandalidae", "Parapaguridae", "Platymaia", 
  "Scyllaridae", "Sergestidae", "Solenoceridae", "Synaxidae", "Nautilus", 
  "Buccinidae", "Astropectinidae", "Goniasteridae", "Isselicrinidae", 
  "Aspidodiadematidae", "Cidaridae", "Diadematidae", "Echinothuriidae", 
  "Elpidiidae", "Euryalidae", "Rhinochimaeridae", "Chimaeridae", "Aulopidae", 
  "Lophotidae", "Ophidiidae", "Epigonidae", "Sternoptychidae", "Stomiidae", 
  "Trachichthyidae", "Halosauridae", "Macrouridae"
)

# ==============================================================================
# 2. QUANTITATIVE METRICS
# ==============================================================================

# A. Calculate depth ranges for ALL taxa
taxon_depths <- clean_taxa %>%
  group_by(phylum, class, order, family, scientificName) %>%
  summarise(
    min_d = min(depth_m, na.rm = TRUE),
    max_d = max(depth_m, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Depth_m = if_else(min_d == max_d, 
                      as.character(min_d), 
                      paste0(min_d, " - ", max_d))
  ) %>%
  select(-min_d, -max_d)

# B. Calculate quantitative stats strictly for ATTRACTED taxa
taxon_quant <- clean_taxa %>%
  filter(attracted == "Attracted") %>%
  group_by(phylum, class, order, family, scientificName) %>%
  summarise(
    freq_occ = round((n_distinct(deployment) / total_deps) * 100),
    MaxN = max(individualCount, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Freq_occ = as.character(freq_occ),
    MaxN = as.character(MaxN)
  ) %>%
  select(-freq_occ)

# C. Extract Tentative IDs (Collapsed if there are multiple; formatting with italics before collapsing)
tentative_list <- clean_taxa %>%
  filter(!is.na(tentativeId)) %>%
  mutate(
    # Use the exact same logic to identify tentative species vs genera
    t_is_species = str_detect(tentativeId, " "),
    t_is_genus = !t_is_species & family != "" & tentativeId != family,
    
    # Apply markdown italics 
    fmt_tentativeId = case_when(
      t_is_species ~ paste0("*", tentativeId, "*"),
      # If you want tentative genera to ALSO have " sp.", use paste0("*", tentativeId, "* sp.") below instead:
      t_is_genus ~ paste0("*", tentativeId, "* sp."), 
      TRUE ~ tentativeId
    )
  ) %>%
  group_by(phylum, class, order, family, scientificName) %>%
  summarise(
    # Collapse the newly formatted IDs!
    t_ids = paste(unique(fmt_tentativeId), collapse = ", "),
    .groups = "drop"
  )

# ==============================================================================
# 3. DATA MERGING & FORMATTING
# ==============================================================================
table_data <- taxon_depths %>%
  left_join(taxon_quant, by = c("phylum", "class", "order", "family", "scientificName")) %>%
  left_join(tentative_list, by = c("phylum", "class", "order", "family", "scientificName")) %>%
  
  mutate(
    # Apply phylogenetic sorting rules
    phylum = factor(phylum, levels = phylo_phyla),
    class = fct_relevel(factor(class), phylo_classes),
    order = fct_relevel(factor(order), phylo_orders),
    family = fct_relevel(factor(family), phylo_families)
  ) %>%
  arrange(phylum, class, order, family, scientificName) %>%
  
  # Fill NA values with blanks
  mutate(across(everything(), as.character)) %>% 
  mutate(across(everything(), ~replace_na(.x, ""))) %>%
  
  # Text Formatting Logic (Italics & Tentative IDs via Markdown)
  mutate(
    # Identify taxonomy levels based on naming conventions
    is_species = str_detect(scientificName, " "),
    is_genus = !is_species & family != "" & scientificName != family,
    
    # Apply base italics and " sp."
    base_name = case_when(
      is_species ~ paste0("*", scientificName, "*"),
      is_genus ~ paste0("*", scientificName, "* sp."),
      TRUE ~ scientificName
    ),
    
    # If a tentative ID exists, append it with a line break <br>
    scientificName = if_else(
      t_ids != "",
      paste0(base_name, "<br>(Tentative: ", t_ids, ")"),
      base_name
    )
  ) %>%
  # Clean up our temporary logic columns
  select(-is_species, -is_genus, -base_name, -t_ids)

# ==============================================================================
# 4. GENERATE GT TABLE
# ==============================================================================
table_data %>%
  gt() %>%
  tab_header(
    title = "Taxonomic Inventory of Deep-Sea Camera Observations",
    subtitle = "Tuvalu"
  ) %>%
  cols_label(
    phylum = "Phylum",
    class = "Class",
    order = "Order",
    family = "Family",
    scientificName = "Lowest Taxonomic ID",
    Depth_m = "Depth (m)",
    Freq_occ = "Freq. of Occ. (%)",
    MaxN = "MaxN"
  ) %>%
  # Center align math columns
  cols_align(
    align = "center",
    columns = c(Depth_m, Freq_occ, MaxN)
  ) %>%
  # Set tight widths
  cols_width(
    Depth_m ~ px(140),  
    Freq_occ ~ px(90),  
    MaxN ~ px(80),
    # Bumped scientificName slightly to accommodate long tentative text
    scientificName ~ px(220) 
  ) %>%
  # TELL GT TO READ THE ASTERISKS AND <br> AS MARKDOWN/HTML
  fmt_markdown(columns = scientificName) %>%
  # Prevent depth column from wrapping
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = Depth_m)
  ) %>%
  opt_row_striping() %>%
  tab_options(
    data_row.padding = px(5),
    table.font.size = px(12)
  )
```


